---
title: "ATAC_csaw"
author: "Abby Deaven"
date: "2025-10-13"
output: html_document
---

```{r echo = TRUE, eval = FALSE }

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager", type = "source")

BiocManager::install("csaw") #INSTALL from source (I had to download a gfortran installer from CRAN (https://cran.r-project.org/bin/macosx/tools/) )

if (!require("BiocManager", quietly = TRUE))
    install.packages("c*BiocManager")

BiocManager::install(c("DiffBind", "ggplot2", "dplyr"))

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("profileplyr")

#edgeR
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("edgeR")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ComplexHeatmap")
BiocManager::install("genomation")
BiocManager::install("stringr")
  library(stringr)

workingdir="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC"

#set working directory to the correct location for working machine
knitr::opts_knit$set(root.dir = workingdir)
setwd("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC")
library(dplyr)
```
Run ChIPQC on ATAC samples
Externally -- make samplesheet with bam file and peak information
```{r}
### ChIPQC ###
install.packages("ChIPQC")
library(dplyr)
library(BiocParallel)

## note - replaced all "-" in filenames with "_" --
# in windows powershell, run: ls *.bam*| Rename-Item -NewName {$_.name -replace "-","_"}
QC_meta <- read.table("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Run154_SampleSheet.txt", sep = "\t", header = TRUE)

#select samples
QC <- QC_meta[1:23,1:8]
# <- loadDb("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Justina/A1163_txdb.sqlite")
read.table("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/SortedBamFiles/152_N15_ATAC_H1mnG_50conidiophore_Rep2.shifted.bam")
BiocParallel::register(BiocParallel::SerialParam())
anaylsis <- ChIPQC(QC, chromosomes=NULL)

metrics <- QCmetrics(anaylsis)
write.table(metrics, file="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/ATAC_QC/QC_metrics.txt", sep="\t")
asdf
ChIPQCreport(anaylsis, reportName="Run152 CEA17 ATAC QC report", reportFolder="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/ATAC_QC", facet = F)

```


#Make an org.db and txdb object for annotation-
```{r}
BiocManager::install(c("AnnotationHub", "AnnotationForge"))
library(stringr)
library(AnnotationForge)

af_gff <- read.table("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/Afum_A1163_edit6_fungidb.gff", sep="\t", quote="")
af_gaf <- read.table("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/FungiDB-CURRENT_AfumigatusA1163_GO.gaf", sep="\t", quote="",skip=1)

af_gff_db <- af_gff %>% dplyr::filter(V3=="gene") 
  af_gff_db$GID <- str_extract(af_gff_db$V9, "AFUB_[0-9]{6}")
#most genes do not have names/symbols in this annotation; copying AFUB number to symbol column. keeping description as description
  af_gff_db$SYMBOL <- str_extract(af_gff_db$V9, "AFUB_[0-9]{6}")
  af_gff_db$GENENAME <- str_extract(af_gff_db$V9,regex("description=.*;"))
    af_gff_db$GENENAME <- gsub("description=","", af_gff_db$GENENAME)
    af_gff_db$GENENAME <- gsub(";.*","", af_gff_db$GENENAME)

gene_info <- af_gff_db %>% dplyr::select(GID, SYMBOL, GENENAME) %>%
  mutate(GENENAME=case_when(is.na(GENENAME) ~ GID,
                           !is.na(GENENAME) ~ GENENAME))

gene_info <- gene_info %>%
  distinct(GID, .keep_all = TRUE) %>%
  filter(!is.na(GID))

af.go <- af_gaf %>% dplyr::filter(V15=="AspGD") %>% dplyr::select(V2, V6, V15) %>% distinct(V2, .keep_all = TRUE)
colnames(af.go) = c("GID", "GO", "EVIDENCE")

#Make dataframe of chromosome sizes
chr.sizes <- data.frame(chrom=c("A1163_Afum_chr1", "A1163_Afum_chr2", "A1163_Afum_chr3", "A1163_Afum_chr4", "A1163_Afum_chr5", "A1163_Afum_chr6", "A1163_Afum_chr7", "A1163_Afum_chr8", "DS499638.1_unaligned", "DS499625.1_unaligned", "DS499613.1_unaligned"), length=c("4929", '4777','4278','3350', '4340', '3746', '1857', '1932', '1248', '2594', '5008'))

gene.chr <- af_gff_db %>% dplyr::select(GID, V1)
colnames(gene.chr) = c("GID", "CHR")

makeOrgPackage(
  gene_info = gene_info, chromosome=gene.chr, go=af.go,
  version = "0.1",
  maintainer = "Abigail Deaven <ad45368@uga.edu>",
  author = "Abigail Deaven",
  outputDir = "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome",
  tax_id = "451804",
  genus = "Aspergillus",
  species = "fumigatus"
)
library(txdbmaker)
Af_txdb <- makeTxDbFromGFF("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/Afum_A1163_edit6_fungidb.gff",
                format='gff',
                dataSource='fungiDB',
                organism="Aspergillus fumigatus",
                taxonomyId=451804,
                circ_seqs=NULL,
                chrominfo=,
                miRBaseBuild=NA,
                metadata=NULL)

saveDb(Nc12, file=)

## load org.db
install.packages("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/org.Afumigatus.eg.db",
                 repos = NULL,
                 type = "source")

library(org.Afumigatus.eg.db)

```

#OLD STEP - Using Diffbind to create consensus peakset
Replaced with bedtools intersect
```{r}
library(DiffBind)
library(dplyr)
library(genomation)
library(GenomicRanges)
samples <- readBed("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Run154_SampleSheet.txt", sep="\t", header=TRUE)
samples <- samples[,1:8]

##add control information

diff_sheet <- samples %>% mutate(ControlID="N24_Genomic_CEA10", bamControl = "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/bamFiles/154_N24_ATAC_gDNA_shifted.marked.dupe.bam", Treatment = "NA", PeakCaller="narrow") %>% relocate(Treatment, .after = Condition)

head(diff_sheet)
diff_sheet <- diff_sheet %>% filter(Tissue != "genomic")
diff_sheet$group = paste(diff_sheet$Tissue, diff_sheet$Condition, sep=".")
lev = factor(diff_sheet$group)
lev = levels(lev)
l='conidia.c37'

for (l in lev) {
print(l)
  
  diff_sheet_l <- diff_sheet %>% filter(group %in% l) %>% select(-group)
  diff_obj_l <- dba(sampleSheet=diff_sheet_l)
  
  nrep=nrow(diff_sheet_l)
  thresh= nrep/(nrep+1)
  
  overlap_rate_l <- dba.overlap(diff_obj_l,mode=DBA_OLAP_RATE)
  
  diff_obj_lconsensus <- dba.peakset(diff_obj_l, consensus= -DBA_REPLICATE, minOverlap=thresh)
  consensus_peaks_l <- dba.peakset(diff_obj_lconsensus, diff_obj_lconsensus$masks$Consensus, bRetrieve=TRUE)
  consensus_peaks_lreduce <- reduce(consensus_peaks_l, drop.empty.ranges=TRUE)

sum(width(consensus_peaks_l))
sum(width(consensus_peaks_lreduce))

  dba.show(diff_obj_lconsensus, mask=diff_obj_lconsensus$masks$Consensus)

  assign(paste0("consensus_",l), consensus_peaks_l)
```

#NEW METHOD - Read in peaksets and merge with GRanges
```{r}
library(genomation)
## Read in merged peaksets to GRanges
Conidia37.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/Conidia37_consensus.bed", sep="\t")

Conidia37.151.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/151_Conidia37_consensus.bed", sep="\t")

Hyphae37.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/151_Hyphae37_consensus.bed", sep="\t")

Conidia50.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/Conidia50_consensus.broadPeak", sep="\t")

Cphore37.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/Conidiophore37_consensus.broadPeak", sep="\t")

Cphore50.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/Conidiophore50_consensus.broadPeak", sep="\t")

Hyphae50.peaks <- readGeneric("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/Hyphae50_Peaks.bed", sep="\t")

## Merge peaks for each tissue -- reduce will concatenate overlapping and nearby peaks so there are not duplicates
  conidia_merge <- reduce(x=c(Conidia37.151.peaks, Conidia37.peaks))
  hyphae_merge <- reduce(x=c(Hyphae37.peaks, Hyphae50.peaks))
  cphore_merge <- reduce(x=c(Cphore37.peaks, Cphore50.peaks))

## Merge ALL peaks
  all_peaks <- reduce(x=c(conidia_merge, hyphae_merge, cphore_merge))
  

```


```{r load files and peaksets}
library(csaw)
#read in sample sheet
samples <- read.table("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Run154_SampleSheet.txt", sep="\t", header=TRUE)
samples <- samples[,1:8]

head(samples)

#remove excluded samples 
exclude=c("154_N28_50_conidia_4", "154_N29_37_conidia_1", "154_N34_37_condiophore_3", "154_N38_50C_hyphae_1")
samples <- samples %>% filter(!SampleID %in% exclude)

bamReads <- samples$bamReads

## extract 3 random bam files to obtain frag sizes
bam1 = samples %>% slice_sample(n=1) %>% pull(bamReads)
bam2 = samples %>% slice_sample(n=1) %>% pull(bamReads)
bam3 = samples %>% filter(Tissue == "conidia") %>% slice_sample(n=1) %>% pull(bamReads)

####QC check: look at fragment sizes of PE data to get max.frag paramenter used above ##1 
 out1 <- getPESizes(bam1)
 frag.sizes1 <- out1$sizes[out1$sizes<= 1500]
 hist(frag.sizes1, breaks=50, xlab="Fragment sizes (bp)", ylab="Frequency", main="", col="grey80")

 ####QC check: look at fragment sizes of PE data to get max.frag paramenter used above ##2
 out2 <- getPESizes(bam2)
 frag.sizes2 <- out2$sizes[out2$sizes<= 1500]
 hist(frag.sizes2, breaks=50, xlab="Fragment sizes (bp)", ylab="Frequency", main="", col="grey80")
 
####QC check: look at fragment sizes of PE data to get max.frag paramenter used above ##3
 out3 <- getPESizes(bam3)
 frag.sizes3 <- out3$sizes[out3$sizes<= 1500]
 hist(frag.sizes3, breaks=50, xlab="Fragment sizes (bp)", ylab="Frequency", main="", col="grey80")


###Example commands to import data from individual experiments
# input <- samples[samples$Antibody == "input"]
# K27samples <- samples[samples$Antibody == "H3K27me3", ]
# K9samples <- samples[samples$Antibody == "H3K9me3", ]
# K36samples <- samples[samples$Antibody == "H3K36me3",]
# Kacsamples <- samples[samples$Antibody == "Kac",]

## Find best fragment size
max.delay <- 500
x <- correlateReads(bamReads, max.delay, param=param)
frag.len = maximizeCcf(x) ### using fragment size 198


# read counts in shared peaks
gdna_peaks <- readGeneric('C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/SamplesForAnalysis/MergePeaks/gDNA_consensus.broadPeak', sep="\t")

param <- readParam(pe="first", max.frag=500, minq=25, discard= gdna_peaks)

peak.counts <- regionCounts(bamReads, all_peaks, param=param)

# MACS2 peaks only: filter low abundance peaks
library("edgeR")
peak.abundances <- aveLogCPM(asDGEList(peak.counts)) 
peak.counts.filt <- peak.counts[peak.abundances > -3, ] # only use peaks logCPM > -3
# few or no peaks should be removed; modify as desired


# #inspect the counts
 head(assay(peak.counts.filt))
# #view the window size locations
 head(rowRanges(peak.counts.filt))
# #view the sample names
# head(colData(data_csaw))
# #preview totals
 peak.counts.filt$totals
 
 ###############################
# count BAM background bins (for TMM normalization)
 bin.size <- 2000L
binned <- windowCounts(bamReads, bin=TRUE, width=10000, param=param)

##########################################
 # method 1: MACS2 peaks only, TMM normalization based on binned counts
peak.counts.tmm <- peak.counts.filt
peak.counts.tmm <- normFactors(binned, se.out=peak.counts.tmm)

```

#Set up differential analysis

```{r}
grouping <- factor(paste(samples$Tissue, samples$Condition, sep="."))

#create a DGElistK9regions object
DGElist <- asDGEList(peak.counts.tmm, group=factor(grouping)) 
                                        #this creates an object with a counts DF and a samples DF

#replace generic "Sample1, Sample2" names with actual names of samples
colnames(DGElist$counts) <- paste(paste(samples$Tissue,samples$Condition, "R", sep="."), samples$Replicate, sep="")
rownames(DGElist$samples) <-  paste(paste(samples$Tissue,samples$Condition, "R", sep="."), samples$Replicate, sep="")

#examine data
png(filename = "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Plots/csaw_mds.png", width = 600, height = 800)
plotMDS(DGElist, method="bcv", col=as.numeric(DGElist$samples$group))
dev.off()
    #plot shows replicates behave well.
d1 <- estimateCommonDisp(DGElist, verbose=T)

#design a model matrix to compare groups
design.mat <- model.matrix(~0 + DGElist$samples$group)
#replace wonky column names with readable ones
colnames(design.mat) <- levels(DGElist$samples$group)

#############################
###
###
#     Estimate Dispersion
###
###
#############################
y <- estimateDisp(DGElist, design.mat)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design.mat, robust=TRUE)
summary(fit$var.post)

head(y$counts)


#plot fit
par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
     ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
     ylab=("Biological coefficient of variation"))
plotQLDisp(fit)
```


## export cpm from edgeR object for heatmapping

# identify all enriched regions in each condition
```{r}
library(txdbmaker)
library(ChIPseeker)

levels = colnames(design.mat)
contrast_gdna= data.frame(sample = levels)

contrast_gdna$contrast <- paste0(levels, "-genomic.gdna")
contrast_gdna <- contrast_gdna[-c(6:7),2]

## no significant enrichment vs. gdna for conidia50, conidiophore37
contrast_differential = c("conidia.c37_151-hyphae.c37_sg", "conidia.c37-hyphae.c37_sg", "conidia.c37-cphore.c37", "conidia.c50-hyphae.c50", "conidia.c50-cphore.c50", "hyphae.c50-hyphae.c37_sg", "conidia.c50-conidia.c37", "cphore.c50-cphore.c37")

    ## no differential peaks: conidia.c50-conidia.c37, cphore.c50-cphore.c37,
#contrasts for testing loop
#contrast="conidia.c37-genomic.control"
#contrast="conidiophore.c50-conidiophore.c37"

contrasts=c(contrast_gdna, contrast_differential)


#CEA10_txdb <- loadDb("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Justina/A1163_txdb.sqlite")

library(ChIPpeakAnno)
library(ChIPseeker)
library(stringr)

contrasts
contrast="conidia.c37-genomic.gdna"

#set up logFC matrix for density plots
score.gdna <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(score.gdna) <- c('name', 'logFC')

score.diff <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(score.diff) <- c('name', 'logFC')



for (contrast in contrasts) {
  contrastname = gsub(".c37","37",contrast)
  contrastname2 = gsub(".c50","50", contrastname)
  contrastname3 = gsub("genomic.gdna","gDNA", contrastname2)
  final_contrastname=gsub("-","_vs_", contrastname3)

  print(final_contrastname)
  
  comparison = makeContrasts(contrasts = contrast, levels=design.mat)
  counts <- peak.counts.tmm
  results_gdna <- glmQLFTest(fit, contrast=comparison)
  rowData(counts) <- cbind(rowData(counts), results_gdna$table)
  merged_gdna <- mergeWindows(rowRanges(counts), tol=100L, max.width=1000L)
  # use most significant window as statistical representation for p-value and FDR for merged windows
  tab.best <- getBestTest(merged_gdna$id, results_gdna$table)
  # concatenating all relevant statistical data for final merged windows (no redundant columns)
  final <- GRanges(cbind(as.data.frame(merged_gdna$region), results_gdna$table[tab.best$rep.test, -4], tab.best[,-c(7:8)]))
  FDR.thresh <- 0.05 # set as desired
#  final <- final[order(final@elementMetadata$FDR), ]
#  final <- final[final@elementMetadata$FDR < FDR.thresh, ]
  ## save foldchange values to new dataframe for downstream analysis
#  final.dataframe <- DataFrame(final)
  
    ## annotate windows
  anno <- detailRanges(
    final,
    txdb = Af_txdb,
    orgdb = org.Afumigatus.eg.db,
    key.field = "SYMBOL",
    promoter = c(1000, 300)
)

mcols(final) <- cbind(mcols(final), anno)
final$gene <- anno$overlap
final$gene <- str_extract(final$gene, "AFUB_[0-9]{6}")

final$feature <- str_extract(anno$overlap, regex(":.:[A-Z]{1}"))
final$feature <- gsub(":.:", "", final$feature)


final.df <- as.data.frame(final)

  assign(paste0("res_", final_contrastname), final)
  
  filename_all <- paste0("./csaw_Analysis/Results/Windows/", final_contrastname, "_all_windows.txt")

  write.table(final, file=filename_all, sep="\t", row.names=FALSE, col.names=TRUE, quote=FALSE)

  #create bedfile for plotting -- will keep different foldchange cutoff for differential vs. genomic comparisons
    if (grepl("gDNA", final_contrastname, fixed = TRUE)) {
      print("gDNA mode")
      
  ## Add to df for heatmapping
    res.df <- final.df %>% dplyr::select(seqnames, start, end, gene, feature, FDR, logFC)
    res.df$region = paste(res.df$seqnames, res.df$start, res.df$end, sep=".")
      res.df.genic <- res.df %>% filter(gene != "" & feature == "P")
      
    # adjust foldchange values based on significance
    res.df <- res.df %>% mutate(log.fdr.adj = case_when(FDR > 0.05 ~ 0.1,
                                                             TRUE ~ logFC))
    new.colname=paste0(final_contrastname,"log.adj")
    res.df <- rename_with(res.df, ~paste0(final_contrastname,".log.adj", recycle0 = TRUE),"log.fdr.adj")
    fdr.res <- res.df[,c(8,9)]
    
      res.df.genic <- res.df.genic %>% mutate(log.fdr.adj = case_when(FDR > 0.05 ~ 0.1,
                                                             TRUE ~ logFC))
      new.colname=paste0(final_contrastname,"log.adj")
      res.df.genic <- rename_with(res.df.genic, ~paste0(final_contrastname,".log.adj", recycle0 = TRUE),"log.fdr.adj")
      fdr.res.genic <- res.df.genic[,c(8,9)]

    ## ON FIRST RUN ONLY - make dataframe for heatmap
          if (grepl("res_conidia37_vs_gDNA", final_contrastname, fixed = TRUE)) {
           print("First run! Making heatmap df")
            
           overlap=final$gene
           overlap=ifelse(is.na(final$gene), ".", final$gene)
           gdna.atac.res.df <-data.frame(seqnames=seqnames(final),
           start=start(final),
           end=end(final),
           names=overlap,
           scores=c(rep(".", length(final))),
           strand=strand(final))
           gdna.atac.res.df$region = paste(gdna.atac.res.df$seqnames, gdna.atac.res.df$start, gdna.atac.res.df$end, sep=".")

           gdna.atac.res.df <- gdna.atac.res.df %>% left_join(fdr.res, by="region")
           
        ## Genic version
           gdna.genic.atac.res.df <- res.df.genic %>% select(seqnames,start,end,gene, region)
           gdna.genic.atac.res.df <- gdna.genic.atac.res.df %>% left_join(fdr.res.genic, by="region")

          }
    ## all other runs - append to data frame
        else { gdna.atac.res.df <- gdna.atac.res.df %>% left_join(fdr.res.genic)
             gdna.genic.atac.res.df <- gdna.genic.atac.res.df %>% left_join(fdr.res.genic, by="region")
             
}
  ## Filter to significant for output      
    final.sig <- final.df %>%
    filter(FDR < 0.05) %>%
    filter(logFC > 0.8)
    
     final.sig.bed <- final.sig %>% 
  dplyr::select(seqnames, start, end, gene, logFC) %>%
  dplyr::rename("chrom" = seqnames) %>%
  dplyr::rename("chromStart" = start) %>%
  dplyr::rename("chromEnd" = end) %>%
  dplyr::rename("Name" = gene) %>%
  dplyr::rename("Score" = logFC) %>%
  mutate(Strand = ".")
     
      bedname_sig <- paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/BedFiles/", final_contrastname, "_sig_windows.bed")
  write.table(final.sig.bed, file=bedname_sig, sep="\t", row.names=FALSE, col.names=FALSE, quote=FALSE)
      assign(paste0("regions_", final_contrastname), final.sig.bed)

      #for density plots, also make a variable with contrast name and foldchange
      final.score <- data.frame(logFC=final.df$logFC)
      final.score$name = rep(final_contrastname, nrow(final.df))
      score.gdna <- rbind(score.gdna, final.sig.score)

      assign(paste0("score_", final_contrastname), final.sig.score)

     
     final.sig.genesonly <- final.sig.bed %>% filter(!is.na(Name))
      assign(paste0("genes_", final_contrastname), final.sig.genesonly)
      
      
      genebedname_sig <- paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/Annotation/", final_contrastname, "_sig_genes.bed")
  write.table(final.sig.genesonly, file=genebedname_sig, sep="\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
    
    }

  ## store genes as a dataframe for downstream analysis

     else {
      print("differential")
    res.df <- final.df %>% dplyr::select(seqnames, start, end, gene, feature, FDR, logFC)
    res.df$region = paste(res.df$seqnames, res.df$start, res.df$end, sep=".")
      res.df.genic <- res.df %>% filter(gene != "" & feature == "P")
      
    # adjust foldchange values based on significance
    res.df <- res.df %>% mutate(log.fdr.adj = case_when(FDR > 0.05 ~ 0.1,
                                                             TRUE ~ logFC))
    new.colname=paste0(final_contrastname,"log.adj")
    res.df <- rename_with(res.df, ~paste0(final_contrastname,".log.adj", recycle0 = TRUE),"log.fdr.adj")
    fdr.res <- res.df[,c(8,9)]
    
      res.df.genic <- res.df.genic %>% mutate(log.fdr.adj = case_when(FDR > 0.05 ~ 0.1,
                                                             TRUE ~ logFC))
      new.colname=paste0(final_contrastname,"log.adj")
      res.df.genic <- rename_with(res.df.genic, ~paste0(final_contrastname,".log.adj", recycle0 = TRUE),"log.fdr.adj")
      fdr.res.genic <- res.df.genic[,c(8,9)]

    ## ON FIRST RUN ONLY - make dataframe for heatmap
          if (grepl("conidia37_151_vs_hyphae37_sg", final_contrastname, fixed = TRUE)) {
           print("First run! Making heatmap df")
            
           overlap=final$gene
           overlap=ifelse(is.na(final$gene), ".", final$gene)
           diff.atac.res.df <-data.frame(seqnames=seqnames(final),
           start=start(final),
           end=end(final),
           names=overlap,
           scores=c(rep(".", length(final))),
           strand=strand(final))
           diff.atac.res.df$region = paste(diff.atac.res.df$seqnames, diff.atac.res.df$start, diff.atac.res.df$end, sep=".")

           diff.atac.res.df <- diff.atac.res.df %>% left_join(fdr.res, by="region")
           
        ## Genic version
           diff.genic.atac.res.df <- res.df.genic %>% select(seqnames,start,end,gene, region)
           diff.genic.atac.res.df <- diff.genic.atac.res.df %>% left_join(fdr.res.genic, by="region")

          }
    ## all other runs - append to data frame
        else { diff.atac.res.df <- diff.atac.res.df %>% left_join(fdr.res.genic)
             diff.genic.atac.res.df <- diff.genic.atac.res.df %>% left_join(fdr.res.genic, by="region")
}

    final.df <- as.data.frame(final)
    final.sig <- final.df %>%
    filter(FDR < 0.05) %>%
      filter(logFC > 1 | logFC < 1)
    
     final.sig.bed <- final.sig %>% 
  dplyr::select(seqnames, start, end, gene, logFC) %>%
  dplyr::rename("chrom" = seqnames) %>%
  dplyr::rename("chromStart" = start) %>%
  dplyr::rename("chromEnd" = end) %>%
  dplyr::rename("Name" = gene) %>%
  dplyr::rename("Score" = logFC) %>%
  mutate(Strand = ".")
     
      bedname_sig <- paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/BedFiles/", final_contrastname, "_sig_windows.bed")
  write.table(final.sig.bed, file=bedname_sig, sep="\t", row.names=FALSE, col.names=TRUE, quote=FALSE)
      assign(paste0("regions_", final_contrastname), final.sig.bed)
      #for density plots, also make a variable with contrast name and foldchange
final.score <- data.frame(logFC=final.df$logFC)
final.score$name = rep(final_contrastname, nrow(final.df))

      assign(paste0("score_", final_contrastname), final.sig.score)
      score.diff <- rbind(score.diff, final.sig.score)
     
     final.sig.genesonly <- final.sig.bed %>% filter(!is.na(Name))
      assign(paste0("genes_", final_contrastname), final.sig.genesonly)
    
genebedname_sig <- paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/Annotation/", final_contrastname, "_sig_genes.bed")
  write.table(final.sig.genesonly, file=genebedname_sig, sep="\t", row.names=FALSE, col.names=FALSE, quote=FALSE)
    }

}
```

# Compare gene accessibility vs. non-accessible genew. HAVE NOT RERUN WITH NEW DATA.
```{r}
### comparing genes in conidia 37 vs. cphore 37
## assigning differential genes to one set or the other
conidia37.v.hyphae37_genes <- data.frame(GeneID = genes_conidia37_151_vs_hyphae37_sg$Name, score=genes_conidia37_151_vs_hyphae37_sg$Score)
conidia37.v.gdna_genes <- data.frame(GeneID = genes_conidia37_151_vs_gDNA$Name)
hyphae37.v.gdna_genes <- data.frame(GeneID = genes_hyphae37_sg_vs_gDNA$Name)

conidia37.v.hyphae37_genes <- conidia37.v.hyphae37_genes %>% 
#  filter(GeneID %in% hyphae37.v.gdna_genes$GeneID) %>%
#  mutate(Identity="Shared") %>% 
  mutate(Differential = case_when(GeneID %in% conidia37.v.hyphae37_genes$GeneID ~ "Yes",
                                  ! GeneID %in% conidia37.v.hyphae37_genes$GeneID ~ "No")) %>%
  mutate(score = as.numeric(score)) %>%
  dplyr::mutate(Conidia.vs.Hyphae = case_when(score > 1 ~ "C>H",
                                      score < 1 ~ "C<H"))

conidia37.v.hyphae37_genes.SHARED <- conidia37.v.gdna_genes %>%
  filter(GeneID %in% hyphae37.v.gdna_genes$GeneID) %>%
  mutate(Identity="Shared")

hyphae37.only <- hyphae37.v.gdna_genes %>% 
  filter(! GeneID %in% conidia37.v.gdna_genes$GeneID) %>% 
  filter(! GeneID %in% conidia37.v.hyphae37_genes.SHARED$GeneID) %>% 
  mutate(Identity="Hyphae37 Only")

conidia37.only <- conidia37.v.gdna_genes %>% 
  filter(! GeneID %in% hyphae37.v.gdna_genes$GeneID) %>% 
  filter(! GeneID %in% conidia37.v.hyphae37_genes.SHARED$GeneID) %>% 
  mutate(Identity="Conidia 37 Only")

All_C37_Genes <- rbind(conidia37.only,hyphae37.only,conidia37.v.hyphae37_genes.SHARED )
All_C37_Genes_diffInfo <- All_C37_Genes %>% left_join(conidia37.v.hyphae37_genes, by = "GeneID") %>%
  mutate(Differential = case_when(is.na(Differential) ~ "No",
                        TRUE ~ Differential)) %>%
  dplyr::select(-score)

conidia <- All_C37_Genes %>% filter(Identity == "Conidia 37 Only" | Identity == "Shared") %>% pull(GeneID)
hyphae <- All_C37_Genes %>% filter(Identity == "Hyphae37 Only" | Identity == "Shared") %>% pull(GeneID)

### Running similar analysis on cphore 50 and cphore 37
Af_genes <- read.table("./Af_Genome/CEA10/A1163_edit6_liftoff.gff", sep="\t", quote="",row.names=NULL, stringsAsFactors = FALSE)
  Af_genes <- Af_genes %>% filter(V3 == "gene") %>% dplyr::select(V1, V4, V5, V9)
  Af_genes$GeneID = str_extract(Af_genes$V9, "AFUB_[0-9]{6}")
  Af_genes <- Af_genes %>% dplyr::select(-V9)

cphore.shared <- data.frame(GeneID = intersect(genes_cphore37_vs_gDNA$geneId, genes_cphore50_vs_gDNA$geneId), ID = "Shared")
cphore.37.only <- data.frame(GeneID= setdiff(genes_cphore37_vs_gDNA$geneId, genes_cphore50_vs_gDNA$geneId), ID = "C37 Only")
cphore.50.only <- data.frame(GeneID= setdiff(genes_cphore50_vs_gDNA$geneId, genes_cphore37_vs_gDNA$geneId), ID = "C50 Only")

cphore_genes <- rbind(cphore.shared, cphore.37.only, cphore.50.only)
cphore_bed <- cphore_genes %>% left_join(Af_genes, by=join_by(GeneID==GeneID)) %>% dplyr::select(-GeneID) %>% relocate(ID, .after=everything())

write.table(cphore_bed, file="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/BedFiles/cphore_genes.bed", sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)
### make venn diagram
install.packages("VennDiagram")
library(VennDiagram)

library(RColorBrewer)
colors <- c("#1a9640", "#da60a0")

## Venn diagram of PRC2-target induction ##
 venn.diagram(
    x = list(conidia, hyphae),
    category.names = c("Conidia", "Hyphae"),
    filename = 'ResultsVenn_Diagram.png',
    output = TRUE,
    imagetype = "png",
    height = 1200,
    width = 1100,
    resolution = 300,
    compression = "lzw",
    lwd = 1,
    lty = 0.5,
    fill = colors,
            cex = .5,
        fontface = "bold",
        fontfamily = "sans",
            cat.cex = 0.5,
        cat.fontface = "bold",
        cat.fontfamily = "sans",
              cat.pos = c(-27, 27),
          cat.dist = c(0.055, 0.055))

write.table(All_C37_Genes_diffInfo, file="./csaw_Analysis/Results/Annotation/All_37C_genes_diffinfo.txt", sep="\t", quote=FALSE, row.names = FALSE)          
conidia.v.hyphae_df <- as.data.frame(res_conidia37_vs_hyphae37)
    conidia.v.hyphae_df.sig <- conidia.v.hyphae_df %>%
    filter(FDR < 0.05) %>%
    filter(logFC > 1 | logFC < -1) %>%
    dplyr::select(seqnames, start, end, width, strand, logFC, FDR, direction) %>%
      mutate(coord=paste0(seqnames,start,end))
    
head(CEA10_txdb)
print(CEA10_txdb)

```

#Density Plots
```{r}
## Make density plots to figure out a cutoff for differential
## First make DF of all logFC values + a sampleID

png(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Plots/FC_over_gDNA_density.png", height=800, width=600)
ggplot(score.gdna, aes(x=logFC, color=factor(name))) + geom_density(alpha=0.5) + xlim(c(-2.5, 4)) + ggtitle("Read Density ATAC") + geom_vline(xintercept = 0.75, linetype="dotted", color="black", size=1) +theme_minimal()
dev.off()

## For inputs -- use a cutoff of logFC > 0.8 to remove background signal

## Differential version
png(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Plots/FC_differential_density.png", height=800, width=600)
ggplot(score.diff, aes(x=logFC, fill=factor(name))) + geom_density(alpha=0.5, color="black") + xlim(c(-2.5, 4)) + ggtitle("Read Density ATAC") + 
  geom_vline(xintercept = 0.8, linetype="dotted", color="blue", size=.75) + 
  geom_vline(xintercept = -0.8, linetype="dotted", color="blue", size=.75) + 
  theme_minimal()
dev.off()

## +/- 0.8 seems like a lenient cutoff; +/-1 is more stringent

```

#Heatmaps
```{r}
atac_Heatmap_df.gdna <- gdna.atac.res.df %>% dplyr::select(region, contains("log.adj"))

### remove any regions that are non-significant in all samples
final_atac_Heatmap_df.gdna <- atac_Heatmap_df.gdna %>% rowwise %>% mutate(rowSum = sum(c_across(abs((2:8)))))
final_atac_Heatmap_df.gdna <- final_atac_Heatmap_df.gdna %>% filter(rowSum != "0.7") %>% select(-rowSum)
#remove outliers/set heatmap scale using whole dataset
write.table(final_atac_Heatmap_df.gdna, "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/ATAC_logFC_vs_gdna_all.txt", sep="\t", quote=FALSE, row.names=FALSE)

atac_Heatmap_mat.gdna <- final_atac_Heatmap_df.gdna[,2:8] 
atac_Heatmap_mat.gdna <- atac_Heatmap_mat.gdna %>% mutate_all(as.numeric)

atac_Heatmap_mat.gdna <- as.matrix(atac_Heatmap_mat.gdna)

library(pheatmap)
library(ComplexHeatmap)
library(RColorBrewer)
library(ggplot2)
library(circlize)

max<- quantile(atac_Heatmap_mat.gdna, 0.99) # identify the 98th percentile value and assign this to max
atac_Heatmap_mat.gdna[atac_Heatmap_mat.gdna>max] =max    #replace any value larger than the 98th percentile value with the 98th percentile value.
min <- quantile(atac_Heatmap_mat.gdna, 0.01)
atac_Heatmap_mat.gdna[atac_Heatmap_mat.gdna<min] = min    

col_fun1<- colorRamp2(c(0, 0.1, quantile(atac_Heatmap_mat.gdna, 0.99)), c("gray96",  "gray96", "darkorchid"))


gdna_hmap <- Heatmap(atac_Heatmap_mat.gdna, col = col_fun1, 
                    cluster_columns = T, show_row_names = FALSE, show_column_dend = FALSE, 
                    show_row_dend = FALSE, cluster_rows = TRUE, 
                    row_title = "Accessible Windows", row_title_side = "left", row_title_gp = gpar(fontsize = 20),
                    heatmap_legend_param = list(col_fun=col_fun1, title = expression("log "[2]~"(Fold Change)"), at = c(0, 1), legend_width = unit(4,"cm"),legend_height=unit(8,'cm'), title_position = "leftcenter-rot", border = "black",labels_gp= gpar(fontsize = 20), title_gp= gpar(fontsize = 20)),
                    column_names_rot = 45, column_names_gp = grid::gpar(fontsize = 14),
                    height = unit(18, 'cm'), width=unit(20, 'cm'), border="black")


   png("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/Plots/ATAC_gdna_heatmap.png", width = 900, height = 800)
   draw(gdna_hmap)
  dev.off()
  
  ## Peaks overlapping promoters
# vs. gDNA

atac_Heatmap_df.genic.gdna <- gdna.genic.atac.res.df %>% dplyr::select(region, gene, contains("log.adj"))

### remove any regions that are non-significant in all samples
atac_Heatmap_df.genic.gdna <- atac_Heatmap_df.genic.gdna %>% rowwise %>% mutate(rowSum = sum(c_across(abs((3:10)))))
final_atac_Heatmap_df.genic.gdna <- atac_Heatmap_df.genic.gdna %>% filter(rowSum != '0.8') %>% select(-rowSum)

write.table(final_atac_Heatmap_df.genic.gdna, "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/ATAC_logFC_vs_gDNA_promoters.txt", sep="\t", quote=FALSE, row.names=FALSE)

#remove outliers/set heatmap scale using whole dataset

final_atac_Heatmap_df.genic.gdna <- final_atac_Heatmap_df.genic.gdna[,3:10] 
final_atac_Heatmap_df.genic.gdna <- final_atac_Heatmap_df.genic.gdna %>% mutate_all(as.numeric)


final_atac_Heatmap_df.genic.gdna <- as.matrix(final_atac_Heatmap_df.genic.gdna)
max3<- quantile(final_atac_Heatmap_df.genic.gdna, 0.99) # identify the 98th percentile value and assign this to max
final_atac_Heatmap_df.genic.gdna[final_atac_Heatmap_df.genic.gdna>max3] =max3 #replace any value larger than the 98th percentile value with the 98th percentile value.
min3 <- quantile(final_atac_Heatmap_df.genic.gdna, 0.01)
final_atac_Heatmap_df.genic.gdna[final_atac_Heatmap_df.genic.gdna<min3] = min3    

col_fun3<- colorRamp2(c(0, 0.1, quantile(final_atac_Heatmap_df.genic.gdna, 0.99)), c("gray96",  "gray96", "darkorchid"))

gdna_genic.hmap <- Heatmap(final_atac_Heatmap_df.genic.gdna, col = col_fun1, 
                    cluster_columns = T, show_row_names = FALSE, show_column_dend = FALSE, 
                    show_row_dend = FALSE, cluster_rows = TRUE, 
                    row_title = "Accessible Promoters", row_title_side = "left", row_title_gp = gpar(fontsize = 20),
                    heatmap_legend_param = list(col_fun=col_fun1, title = expression("log "[2]~"(Fold Change)"), at = c(0, 2), legend_width = unit(4,"cm"),legend_height=unit(8,'cm'), title_position = "leftcenter-rot", border = "black",labels_gp= gpar(fontsize = 20), title_gp= gpar(fontsize = 20)),
                    column_names_rot = 45, column_names_gp = grid::gpar(fontsize = 14),
                    height = unit(18, 'cm'), width=unit(20, 'cm'), border="black")


   png("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/Plots/ATAC_gdna_promoters_heatmap.png", width = 900, height = 800)
   draw(gdna_hmap)
  dev.off()

```
# Make a heatmap for differential samples
```{r}


atac_Heatmap_df.diff <- diff.atac.res.df %>% dplyr::select(region, contains("log.adj"))

### remove any regions that are non-significant in all samples
final_atac_Heatmap_df.diff <- atac_Heatmap_df.diff %>% rowwise %>% mutate(rowSum = sum(c_across(abs((2:9)))))
final_atac_Heatmap_df.diff <- final_atac_Heatmap_df.diff %>% filter(rowSum != '0.8') %>% select(-rowSum)

write.table(final_atac_Heatmap_df.diff, "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/ATAC_logFC_vs_differential_all.txt", sep="\t", quote=FALSE, row.names=FALSE)

#remove outliers/set heatmap scale using whole dataset

final_atac_Heatmap_df.diff <- final_atac_Heatmap_df.diff[,2:9] 
final_atac_Heatmap_df.diff <- final_atac_Heatmap_df.diff %>% mutate_all(as.numeric)

final_atac_Heatmap_df.diff <- as.matrix(final_atac_Heatmap_df.diff)

library(pheatmap)
library(ComplexHeatmap)
library(RColorBrewer)
library(ggplot2)
library(circlize)

max<- quantile(final_atac_Heatmap_df.diff, 0.99) # identify the 98th percentile value and assign this to max
final_atac_Heatmap_df.diff[final_atac_Heatmap_df.diff>max] =max    #replace any value larger than the 98th percentile value with the 98th percentile value.
min <- quantile(final_atac_Heatmap_df.diff, 0.01)
final_atac_Heatmap_df.diff[final_atac_Heatmap_df.diff<min] = min    

col_fun2<- colorRamp2(c(quantile(final_atac_Heatmap_df.diff, 0.01), -0.1, 0, 0.1, quantile(final_atac_Heatmap_df.diff, 0.99)), c("#6DBC90FF","gray96",  "gray96", "gray96", "#C1447EFF"))


diff.hmap <- Heatmap(final_atac_Heatmap_df.diff, col = col_fun2, 
                    cluster_columns = T, show_row_names = FALSE, show_column_dend = FALSE, 
                    show_row_dend = FALSE, cluster_rows = TRUE, 
                    row_title = "Accessible Windows", row_title_side = "left", row_title_gp = gpar(fontsize = 20),
                    heatmap_legend_param = list(col_fun=col_fun2, title = expression("log "[2]~"(Fold Change)"), at = c(-2,0, 2), legend_width = unit(4,"cm"),legend_height=unit(8,'cm'), title_position = "leftcenter-rot", border = "black",labels_gp= gpar(fontsize = 20), title_gp= gpar(fontsize = 20)),
                    column_names_rot = 45, column_names_gp = grid::gpar(fontsize = 14),
                    height = unit(18, 'cm'), width=unit(20, 'cm'), border="black")


   png("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/Plots/ATAC_differential_heatmap.png", width = 900, height = 800)
   draw(diff.hmap)
  dev.off()


draw(diff.hmap)

## Peaks overlapping promoters

atac_Heatmap_df.genic.diff <- diff.genic.atac.res.df %>% dplyr::select(region, gene, contains("log.adj"))

### remove any regions that are non-significant in all samples
atac_Heatmap_df.genic.diff <- atac_Heatmap_df.genic.diff %>% rowwise %>% mutate(rowSum = sum(c_across(abs((3:10)))))
final_atac_Heatmap_df.genic.diff <- atac_Heatmap_df.genic.diff %>% filter(rowSum != '0.8') %>% select(-rowSum)

write.table(final_atac_Heatmap_df.genic.diff, "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/ATAC_logFC_vs_diff_promoters.txt", sep="\t", quote=FALSE, row.names=FALSE)

#remove outliers/set heatmap scale using whole dataset

final_atac_Heatmap_df.genic.diff <- final_atac_Heatmap_df.genic.diff[,3:10] 
final_atac_Heatmap_df.genic.diff <- final_atac_Heatmap_df.genic.diff %>% mutate_all(as.numeric)


final_atac_Heatmap_df.genic.diff <- as.matrix(final_atac_Heatmap_df.genic.diff)
max4<- quantile(final_atac_Heatmap_df.genic.diff, 0.99) # identify the 98th percentile value and assign this to max
final_atac_Heatmap_df.genic.diff[final_atac_Heatmap_df.genic.diff>max4] =max4 #replace any value larger than the 98th percentile value with the 98th percentile value.
min4 <- quantile(final_atac_Heatmap_df.genic.diff, 0.01)
final_atac_Heatmap_df.genic.diff[final_atac_Heatmap_df.genic.diff<min4] = min4    

col_fun4 <-  colorRamp2(c(quantile(final_atac_Heatmap_df.genic.diff, 0.01), -0.1, 0, 0.1, quantile(final_atac_Heatmap_df.genic.diff, 0.99)), c("#6DBC90FF","gray96",  "gray96", "gray96", "#C1447EFF"))

diff_genic.hmap <- Heatmap(final_atac_Heatmap_df.genic.diff, col = col_fun4, 
                    cluster_columns = T, show_row_names = FALSE, show_column_dend = FALSE, 
                    show_row_dend = FALSE, cluster_rows = TRUE, 
                    row_title = "Accessible Promoters", row_title_side = "left", row_title_gp = gpar(fontsize = 20),
                    heatmap_legend_param = list(col_fun=col_fun1, title = expression("log "[2]~"(Fold Change)"), at = c(-2, 0, 2), legend_width = unit(4,"cm"),legend_height=unit(8,'cm'), title_position = "leftcenter-rot", border = "black",labels_gp= gpar(fontsize = 20), title_gp= gpar(fontsize = 20)),
                    column_names_rot = 45, column_names_gp = grid::gpar(fontsize = 14),
                    height = unit(18, 'cm'), width=unit(20, 'cm'), border="black")


   png("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/csaw_Analysis/Results/Plots/ATAC_diff_promoters_heatmap.png", width = 900, height = 800)
   draw(diff_genic.hmap)
  dev.off()


```
## Heatmapping - old version!!
```{r}
draw(gdna_hmap)

## Use foldchange vs. gDNA results to plot a heatmap over all of the regions used for the analysis
## plot everything vs. gDNA
atac.res.df <-data.frame(seqnames=seqnames(res_conidia37_vs_gDNA),
  start=start(res_conidia37_vs_gDNA),
  end=end(res_conidia37_vs_gDNA),
  names=c(rep(".", length(res_conidia37_vs_gDNA))),
  scores=c(rep(".", length(res_conidia37_vs_gDNA))),
  strand=strand(res_conidia37_vs_gDNA),
   logFC_conidia37_vs_gDNA = res_conidia37_vs_gDNA$logFC,
   PValue_conidia37_vs_gDNA = res_conidia37_vs_gDNA$PValue,
   FDR_conidia37_vs_gDNA = res_conidia37_vs_gDNA$FDR,
     logFC_conidia37_151_vs_gDNA = res_conidia37_151_vs_gDNA$logFC,
   PValue_conidia37_151_vs_gDNA = res_conidia37_151_vs_gDNA$PValue,
   FDR_conidia37_151_vs_gDNA = res_conidia37_151_vs_gDNA$FDR,
     logFC_hyphae37_sg_vs_gDNA = res_hyphae37_sg_vs_gDNA$logFC,
     PValue_hyphae37_sg_vs_gDNA = res_hyphae37_sg_vs_gDNA$PValue,
     FDR_hyphae37_sg_vs_gDNA = res_hyphae37_sg_vs_gDNA$FDR,
       logFC_cphore37_vs_gDNA = res_cphore37_vs_gDNA$logFC,
       PValue_cphore37_vs_gDNA = res_cphore37_vs_gDNA$PValue,
       FDR_cphore37_vs_gDNA = res_cphore37_vs_gDNA$FDR,
   logFC_conidia50_vs_gDNA = res_conidia50_vs_gDNA$logFC,
   PValue_conidia50_vs_gDNA = res_conidia50_vs_gDNA$PValue,
   FDR_conidia50_vs_gDNA = res_conidia50_vs_gDNA$FDR,
     logFC_hyphae50_vs_gDNA = res_hyphae50_vs_gDNA$logFC,
     PValue_hyphae50_vs_gDNA = res_hyphae50_vs_gDNA$PValue,
     FDR_hyphae50_vs_gDNA = res_hyphae50_vs_gDNA$FDR,
       logFC_cphore50_vs_gDNA = res_cphore50_vs_gDNA$logFC,
       PValue_cphore50_vs_gDNA = res_cphore50_vs_gDNA$PValue,
       FDR_cphore50_vs_gDNA = res_cphore50_vs_gDNA$FDR)

## make log.adj column --> set between -1 and 1 to 0, then filter out any with sum 0
## then make dataset and remove any that are nonsig in all samples      

atac.res.df_sig  <- within(atac.res.df, logFC_conidia37_vs_gDNA[FDR_conidia37_vs_gDNA > 0.05] <- 0.1)
atac.res.df_sig  <- within(atac.res.df, logFC_conidia37_151_vs_gDNA[FDR_conidia37_151_vs_gDNA > 0.05] <- 0.1)
atac.res.df_sig  <- within(atac.res.df_sig, logFC_hyphae37_sg_vs_gDNA[FDR_conidia37_vs_gDNA > 0.05] <- 0.1)
atac.res.df_sig  <- within(atac.res.df_sig, logFC_cphore37_vs_gDNA[FDR_cphore37_vs_gDNA > 0.05] <- 0.1)
atac.res.df_sig  <- within(atac.res.df, logFC_conidia50_vs_gDNA[FDR_conidia50_vs_gDNA > 0.05] <- 0.1)
atac.res.df_sig  <- within(atac.res.df_sig, logFC_hyphae50_vs_gDNA[FDR_conidia50_vs_gDNA > 0.05] <- 0.1)
atac.res.df_sig  <- within(atac.res.df_sig, logFC_cphore50_vs_gDNA[FDR_cphore50_vs_gDNA > 0.05] <- 0.1)

atac.res.df_sig2 <- atac.res.df_sig %>% 
  mutate(conidia37_log.adj = case_when(logFC_conidia37_vs_gDNA < 0 & logFC_conidia37_vs_gDNA > - 0.5~ 0.1,
                                       logFC_conidia37_vs_gDNA > 0 & logFC_conidia37_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_conidia37_vs_gDNA
                                       )) %>%
  mutate(conidia37_151_log.adj = case_when(logFC_conidia37_151_vs_gDNA < 0 & logFC_conidia37_151_vs_gDNA > - 0.5~ 0.1,
                                       logFC_conidia37_151_vs_gDNA > 0 & logFC_conidia37_151_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_conidia37_151_vs_gDNA
                                       )) %>%

    mutate(hyphae37_log.adj = case_when(logFC_hyphae37_sg_vs_gDNA < 0 & logFC_hyphae37_sg_vs_gDNA > - 0.5~ 0.1,
                                       logFC_hyphae37_sg_vs_gDNA > 0 & logFC_hyphae37_sg_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_hyphae37_sg_vs_gDNA
                                       )) %>%
  mutate(cphore37_log.adj = case_when(logFC_cphore37_vs_gDNA < 0 & logFC_cphore37_vs_gDNA > - 0.5~ 0.1,
                                       logFC_cphore37_vs_gDNA > 0 & logFC_cphore37_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_cphore37_vs_gDNA
                                       )) %>%
    mutate(conidia50_log.adj = case_when(logFC_conidia50_vs_gDNA < 0 & logFC_conidia50_vs_gDNA > - 0.5~ 0.1,
                                       logFC_conidia50_vs_gDNA > 0 & logFC_conidia50_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_conidia50_vs_gDNA
                                       )) %>%
      mutate(hyphae50_log.adj = case_when(logFC_hyphae50_vs_gDNA < 0 & logFC_hyphae50_vs_gDNA > - 0.5~ 0.1,
                                       logFC_hyphae50_vs_gDNA > 0 & logFC_hyphae50_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_hyphae50_vs_gDNA
                                       )) %>%
  mutate(cphore50_log.adj = case_when(logFC_cphore50_vs_gDNA < 0 & logFC_cphore50_vs_gDNA > - 0.5~ 0.1,
                                       logFC_cphore50_vs_gDNA > 0 & logFC_cphore50_vs_gDNA <  0.5~ 0.1,
                                       TRUE ~ logFC_cphore50_vs_gDNA
                                       ))
rownames(atac.res.df_sig2) <- paste(atac.res.df_sig2$seqnames, atac.res.df_sig2$start, atac.res.df_sig2$end, sep=".")

atac_Heatmap_df <- atac.res.df_sig2 %>% 
  rownames_to_column(var="id.log.adj") %>%
  dplyr::select(contains("log.adj")) %>%
  rowwise() %>%
    mutate(Sum = sum(c_across(2:8))) %>%
    filter(Sum > "0.7") %>%
  column_to_rownames(var="id.log.adj") %>%
  dplyr::select(-Sum)


#atac_Heatmap_df_fdr <- atac.res.df %>% dplyr::select(contains("FDR"))
#rownames(atac_Heatmap_df_fdr) = paste(atac.res.df_sig$seqnames, atac.res.df_sig$start, atac.res.df_sig$end, sep=".")

atac_Heatmap_df <- atac.res.df_sig2 %>% dplyr::select(contains("log.adj"))
rownames(atac_Heatmap_df) <- paste(atac.res.df_sig2$seqnames, atac.res.df_sig2$start, atac.res.df_sig2$end, sep=".")


### remove any regions that are non-significant in all samples
#final_atac_Heatmap_df <- atac_Heatmap_df %>% rowwise %>% mutate(rowSum = sum(c_across(1:3))) %>% filter(rowSum > 0.3)
#final_atac_Heatmap_df<-final_atac_Heatmap_df[,1:3]

library(pheatmap)
library(ComplexHeatmap)
library(RColorBrewer)
library(ggplot2)

#remove outliers/set heatmap scale
## splitting samples vs. gdna and differential samples into 2 separate lists for heatmaps
atac_Heatmap_mat <- as.matrix(atac_Heatmap_df)

max<- quantile(atac_Heatmap_mat, 0.99) # identify the 98th percentile value and assign this to max
atac_Heatmap_mat[atac_Heatmap_mat>max] =max    #replace any value larger than the 98th percentile value with the 98th percentile value.
min <- quantile(atac_Heatmap_mat, 0.01)
atac_Heatmap_mat[atac_Heatmap_mat<min] = min    

col_fun1<- colorRamp2(c(0, 0.1, quantile(atac_Heatmap_mat, 0.99)), c("gray96",  "gray96", "darkorchid"))
col_fun(seq(0, quantile(gdna, 0.99), length = 20))

leg1 = Legend(col_fun= col_fun1, title = expression(bold("log "[2]~"(Fold Change)")), direction= "horizontal", at = c(-2, 0, 2), legend_width = unit(4,"cm"), title_position = "topcenter", border = "black")

gdna_hmap <- Heatmap(atac_Heatmap_mat, col = col_fun1, 
                    cluster_columns = F, show_row_names = FALSE, show_column_dend = FALSE, 
                    show_row_dend = FALSE, cluster_rows = TRUE, 
                    row_title = "Accessible Genes (n=4238)", row_title_side = "left", row_title_gp = gpar(fontsize = 16),
                    heatmap_legend_param = list(col_fun= col_fun1, title = expression(bold("log "[2]~"(Fold Change)")), direction= "horizontal", at = c(0, 2), legend_width = unit(4,"cm"), title_position = "topcenter", border = "black"),
                    column_names_rot = 45, column_names_gp = grid::gpar(fontsize = 14))
#                    column_labels = c("Conidia", "Hyphae"))
gmap = draw(gdna_hmap)
   gdna_hmap_order <- row_order(gmap)
   gdna_hmap_ordernames <- rownames(gdna)[unlist(gdna_hmap_order)]
     
gdna_hmap   
   
                  
min(diff)
min(gdna)

col_fun2 <- colorRamp2(c(-2,0, 0.1, quantile(diff, 0.99)), c("#1a9640", "gray96", "gray96",  "#da60a0"))
col_fun(seq(0, quantile(diff, 0.99), length = 20))
leg2 = Legend(col_fun= col_fun2, title = expression(bold("log "[2]~"(Fold Change)")), direction= "horizontal", at = c(-2, 0, 2), legend_width = unit(4,"cm"), title_position = "topcenter", border = "black")

diff_hmap <- Heatmap(diff, col = col_fun2, 
                    cluster_columns = F, show_row_names = FALSE, show_column_dend = FALSE, 
                    show_row_dend = FALSE, cluster_rows = FALSE, row_order = gdna_hmap_ordernames,
                    row_title = "Accessible Genes (n=4238)", row_title_side = "left", row_title_gp = gpar(fontsize = 16),
                    heatmap_legend_param = list(title = expression(bold("log "[2]~"(Fold Change)")), at = c(0, 2), legend_width = unit(4,"cm"), title_position = "topcenter", border = "black"),
                    column_names_rot = 45, column_names_gp = grid::gpar(fontsize = 14),
                    column_labels = c("Conidia vs Hyphae"))
draw(diff_hmap, heatmap_legend_side = "top")
draw(diff_hmap)
png("./csaw_Analysis/Results/Accessibility_heatmap.png", width = 600, height = 800)
draw(gdna_hmap,  heatmap_legend_side = "left")
dev.off()

library(circlize)
col_fun<- colorRamp2(c(-2, -1, 0.2, quantile(gdna, 0.99)), c("goldenrod2", "gray96",  "gray96", "darkorchid"))
col_fun(seq(-2, quantile(atac_heatmap_mat_out, 0.99), length = 20))




## use final gene set to annotate consensus genes
consensus_peaks_anno <- as.data.frame(consensus_peaks)
consensus_peaks_anno <- consensus_peaks_anno %>%
  dplyr::select(seqnames, start, end) %>%
  dplyr::rename("chrom" = seqnames) %>%
  dplyr::rename("chromStart" = start) %>%
  dplyr::rename("chromEnd" = end)

  write.table(consensus_peaks_anno, file="./csaw_Analysis/Results/BedFiles/consensus_peaks.bed", sep="\t", row.names=FALSE, col.names=FALSE, quote=FALSE)
  
  ## annotate windows from bed files
  peakfile <- readPeakFile("./csaw_Analysis/Results/BedFiles/consensus_peaks.bed")
  consensus_peakAnno <- annotatePeak(peakfile, tssRegion=c(-1000, 500),
                         TxDb=CEA10_txdb)

  consensus_peakAnno <-as.data.frame(consensus_peakAnno)
  consensus_peakAnno_out <- consensus_peakAnno %>% filter(distanceToTSS < 2000 & distanceToTSS > -2000) %>% filter(annotation == "Promoter")
  consensus_peakAnno_out_bed <- consensus_peakAnno_out %>% dplyr::select(seqnames, geneStart, geneEnd, geneId)

  
  
  write.table(consensus_peakAnno_out, file="./csaw_Analysis/Results/Annotation/consensus_genes.txt", sep="\t", row.names=FALSE, col.names=FALSE, quote=FALSE)
    write.table(consensus_peakAnno_out_bed, file="./csaw_Analysis/Results/Annotation/consensus_genes.bed", sep="\t", row.names=FALSE, col.names=FALSE, quote=FALSE)

  ## store genes as a dataframe for downstream analysis


```




```{r}
### Go term analysis
r
### Read in lists of GO terms and enrichment from FungiDB
conidia_bio <- data.frame(read.table("./csaw_Analysis/Results/conidia_bioprocess_go.tsv", sep = "\t", header=TRUE))

  conidia_bio_df <- conidia_bio %>%
    dplyr::select(-Result.gene.list) %>%
    dplyr::rename("go" = ID) %>%
    dplyr::rename("FDR"= "Benjamini") %>%
    dplyr::rename("PercentBackground" = Pct.of.bgd) %>%
    dplyr::arrange(FDR, Fold.enrichment) %>%
    mutate(Tissue = "Conidia")
  conidia_bio_top10 <- head(conidia_bio_df, 10)

hyphae_bio_go <- data.frame(read.table("./csaw_Analysis/Results/hyphae_bioprocess_go.tsv", sep = "\t", header=TRUE))

  hyphae_bio_go_df <- hyphae_bio_go %>%
    dplyr::select(-Result.gene.list) %>%
    dplyr::rename("go" = ID) %>%
    dplyr::rename("FDR"= "Benjamini") %>%
    dplyr::rename("PercentBackground" = Pct.of.bgd) %>%
    dplyr::arrange(FDR, Fold.enrichment) %>%
    dplyr::mutate(Tissue = "Hyphae")
  hyphae_bio_top10 <- head(hyphae_bio_go_df,10)

shared_GO <- data.frame(read.table("./csaw_Analysis/Results/hyphae_bioprocess_go.tsv", sep = "\t", header=TRUE))

  shared_GO_df <- shared_GO %>%
    dplyr::select(-Result.gene.list) %>%
    dplyr::rename("go" = ID) %>%
    dplyr::rename("FDR"= "Benjamini") %>%
    dplyr::rename("PercentBackground" = Pct.of.bgd) %>%
    dplyr::arrange(FDR, Fold.enrichment) %>%
    dplyr::mutate(Tissue = "Shared")
  shared_GO_top10 <- head(shared_GO_df,10)

  
all_top_go <- as.data.frame(rbind(conidia_bio_top10, hyphae_bio_top10, shared_GO_top10))
all_go <- as.data.frame(rbind(conidia_bio_df, hyphae_bio_go_df, shared_GO_df))

write.table(all_go, file="./Supplement/TableS4_FP_Genes_All_GOterms.txt", sep="\t", quote=FALSE)

### Shortening names of GO terms for plotting ###
all_top_go$Rename <- all_top_go$Name
renameframe = all_top_go[,c(2, 12)]

renameframe
all_top_go$Rename <- gsub("olic process","olism", all_top_go$Rename)
all_top_go$Rename <- gsub("etic process","esis", all_top_go$Rename)
all_top_go$Rename <- gsub("nucleic acid-templated","", all_top_go$Rename)

##Wrapping names to improve visuals
all_top_go$NameWrap <- str_wrap(all_top_go$Rename, width=30)

write.table(all_top_go, file="./Supplement/top_goterms.txt", sep="\t")

##Plot as a lollipop plot
GO_lollipop <- ggplot(all_top_go, aes(x = reorder(NameWrap, -log10(FDR)), y = -log10(FDR), color = FDR, label = paste("n =", PercentBackground))) +
  geom_segment(aes(xend = reorder(NameWrap, -log10(FDR)), yend = 0), color = "gray", width=1.25) +
  geom_point(aes(size = PercentBackground)) +
  scale_size_continuous(name=str_wrap("Percent Background", width=8), range = c(10,15)) +
  scale_color_gradient(low = "#5e3c99", high = "#e66101") +
  facet_grid(rows = vars(Tissue), scales="free") +
  coord_flip() +
  labs(x = "GO Term", y = "Fold Enrichment") +
  theme_classic(base_size = 22) +
  theme(
    axis.text.x = element_text(hjust = 1, size = 24),
    axis.text.y = element_text(size = 24),
    axis.title = element_text(size=24),
    legend.text = element_text(size=22),
    plot.background = element_rect(fill = "white"),
    axis.line = element_line(color = "black",
                                 linewidth = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 24))
  )

GO_lollipop
ggsave("./csaw_analysis/Results/GO_plot.png", plot=GO_lollipop, dpi=400, width=50, height=50, units="cm")

```
```{r}
### plot accessibility at conidia genes
install.packages("devtools")
library(devtools)
BiocManager::install("GenomicRanges")
BiocManager::install("rtracklayer")
BiocManager::install("patchwork")

install.packages(c('GenomicRanges', 'rtracklayer'))

devtools::install_github('abbydeaven/TrackPlotR_mod')
library(TrackPlotR)
#source("https://github.com/PoisonAlien/trackplot/blob/master/R/trackplot.R?raw=true")

library(ggplot2)
library(rtracklayer)
library(trackplot)
library(GenomicRanges)
library(dplyr)
library(ggplot2)
library(patchwork)


genome = "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/CEA10/A1163_edit6_liftoff.gff"
af_anno <- rtracklayer::import(con = genome,format = 'gff')
af_anno$type <- gsub("mRNA", "transcript", af_anno$type)

mcols(af_anno)$transcript_id = af_anno$locus_tag
mcols(af_anno)$gene_id = af_anno$locus_tag

path="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/"
bw =paste0(path, c("151_ATAC_CEA10_WT_Hyphae37_marked.min30.max200.bin15.smooth45.bw", 
                        "151_ATAC_CEA10_WT_Conidia37_marked.min30.max200.bin15.smooth45.bw"))
bw_files <- load_BigWig(file_path=bw, file_name=c("Hyphae 37", "Conidia 37"))


hic_genes = data.frame(Name = c("HicB", "HicC", "HicD", "HicF", "HicG", "HicH", "HicJ"), ID = c("AFUB_101290", "AFUB_015790", "AFUB_045200", "AFUB_033900", "AFUB_029950", "AFUB_086630", "AFUB_077980"))

hic_ids = hic_genes$ID

for (h in hic_ids) {
  name = hic_genes %>% filter(ID == h) %>% pull(Name)
  print(paste0(h, "==", name))
  
  atac_track <- coverage_vis_gene(coverage_table = bw_files,gene_anno = af_anno,column_name = 'gene_id',gene = h,up_extend = 1000,down_extend = 1000,style="style_2",y_lim=12, sample_order = c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) + labs(title = name) +     
  theme(axis.ticks.x = element_blank(),                                                                             axis.text.x = element_blank(),                                                                              axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA, color="black"),
        plot.margin = unit(c(0,0,0,0), "cm"),
        plot.title = element_text(hjust = 0.5, size = 16))

print(atac_track)
  gene_track <- gene_track_vis(gene_anno = af_anno,column_name = 'gene_id',gene = h,style="style_2", up_extend = 1000, down_extend = 1000, display_mode = "collapse") + 
  theme(axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())
  
  hic_plot <- plot(atac_track + gene_track +
       plot_layout(ncol=1, heights=c(0.4, 0.1)))

  
  plot_path <- paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/", name, "_trackplot.png")
  ggsave(plot_path, hic_plot, device="png", width=8, height=3, dpi=600)

}

atac <- coverage_vis_gene(coverage_table = bw_files,gene_anno = af_anno,column_name = 'gene_id',gene = 'AFUB_101290',up_extend = 1000,down_extend = 1000,style="style_2",y_lim=12, sample_order = c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) +      
  theme(axis.ticks.x = element_blank(),                                                                             axis.text.x = element_blank(),                                                                              axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA, color="black"))
        plot.margin = unit(c(0,0,0,0), "cm"),
        plot.title = element_text(hjust = 0.5, size = 2))

print(p1 + theme(aspect.ratio = 0.1))

gene <- gene_track_vis(gene_anno = af_anno,column_name = 'gene_id',gene = 'AFUB_101290',style="style_2", up_extend = 1000, down_extend = 1000, display_mode = "collapse") + 
  theme(axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())

hic_plot <- plot(atac + gene +
       plot_layout(ncol=1, heights=c(0.4, 0.2)), guides = "collect")
ggsave("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/hicC_plot.png", hic_plot, device="png", width=8, height=2, dpi=600)
```


```{r}
wholeview = coverage_vis_region(bw_files, whole_view, y_lim=40, y_label = "Coverage", style="style_2", sample_order=c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) +      
  theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                             axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
                                      strip.background = element_rect(fill="white"),
                                      panel.border=element_rect(fill=NA, color="black"))
plot(wholeview)

whole_genes = transcript_vis_region(gene_anno = af_anno,region = whole_view, display_by = 'gene',show_name = NULL, display_mode = "collapse") + theme(        axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())


wholeplot = plot(wholeview + plot_spacer() + whole_genes +
       plot_layout(ncol=1, heights=c(0.5, -.05, 0.05)))

ggsave(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/Chrom3plot.png", plot=wholeplot, device="png", width=12, height=4)

conidia_only = as("A1163_Afum_chr3:1118000-1123000", "GRanges")
  conview = coverage_vis_region(bw_files, conidia_only, y_lim=7, y_label = "Coverage", style="style_2", sample_order=c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) +      
  theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                             axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
                                      strip.background = element_rect(fill="white"),
                                      panel.border=element_rect(fill=NA, color="black"))
         conview = highlight_region(conview, start = "112000", end = "112500")

con_genes = transcript_vis_region(gene_anno = af_anno,region = conidia_only, display_by = 'gene',show_name = "gene_id", display_mode = "collapse") + theme(axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())

conplot <- plot(conview + plot_spacer() + con_genes +
       plot_layout(ncol=1, heights=c(0.4, -.0446, 0.05)))
ggsave("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/conidiaplot.png", conplot, device="png", width=6, height=4, dpi=600)


hyphae_only = as("A1163_Afum_chr3:95500-100500", "GRanges")
 hyview = coverage_vis_region(bw_files, hyphae_only, y_lim=20, y_label = "Coverage", style="style_2", sample_order=c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) +      
  theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                             axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
                                      strip.background = element_rect(fill="white", color= "white"),
                                      panel.border=element_rect(fill=NA, color="black"))
         conview = highlight_region(conview, start = "112000", end = "112500")

hgenes = transcript_vis_region(gene_anno = af_anno,region = hyphae_only, display_by = 'gene',show_name = "gene_id", display_mode = "collapse") + theme(              axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())

hyphae <- plot(hyview + plot_spacer() + hgenes +
plot_layout(ncol=1, heights=c(0.4, -.0446, 0.05)))

ggsave("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/hypaheplot.png", hyphae, device="png", width=6, height=4, dpi=600)

shared = as("A1163_Afum_chr3:420000-425000", "GRanges")
 shareview = coverage_vis_region(bw_files, shared, y_lim=10, y_label = "Coverage", style="style_2", sample_order=c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) +      
  theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                             axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
                                      strip.background = element_rect(fill="white"),
                                      panel.border=element_rect(fill=NA, color="black"))
         conview = highlight_region(conview, start = "112000", end = "112500")

sharegenes = transcript_vis_region(gene_anno = af_anno,region = shared, display_by = 'gene',show_name = "gene_id", display_mode = "collapse") + theme(                     axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())


sharedplot <- plot(shareview + plot_spacer() + sharegenes +
       plot_layout(ncol=1, heights=c(0.4, -.0446, 0.05)))
ggsave("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/shareplot.png", sharedplot, device="png", width=6, height=4, dpi=600)


#### safe haven region
haven = as("A1163_Afum_chr7:1811000-1817500", "GRanges")
 havenplot = coverage_vis_region(bw_files, shared, y_lim=20, y_label = "Coverage", style="style_2", sample_order=c("Hyphae 37", "Conidia 37"), col_pal=c("darkorchid","springgreen4")) +      
  theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                             axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
                                      strip.background = element_rect(fill="white"),
                                      panel.border=element_rect(fill=NA, color="black"))
         conview = highlight_region(conview, start = "112000", end = "112500")

         havengenes = transcript_vis_region(gene_anno = af_anno,region = haven, display_by = 'gene',show_name = "gene_id", display_mode = "collapse") + theme(                     axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank())


haven_plot <- plot(havenplot + plot_spacer() + havengenes +
       plot_layout(ncol=1, heights=c(0.4, -.0446, 0.05)))
ggsave("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/safe_haven_plot.png", haven_plot, device="png", width=8, height=2, dpi=600)


```

plot accessibility at starship genes
```{r}


install.packages("C:/Users/abbyd/bin/TrackPlotR_mod", repos=NULL, type="source")

library(ggplot2)
library(rtracklayer)
library(TrackPlotR)
library(dplyr)
library(GenomicRanges)
library(patchwork)
starships <- read.table("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Starships.bed", sep="\t")
starships_genes <- read.table("../F32_PreliminaryData/starship_genes.bed", sep="\t")

starship=as("A1163_Afum_chr1:561788-579157", "GRanges")
chrom="A1163_Afum_chr1"
region_start="562788"
region_end="58157"

##starship window
window <- GRanges(seqnames = chrom, ranges=IRanges(start = 561788, end = 579157))
t2=transcript_vis_region(gene_anno = genome, )

colnames(starships) = c("seqnames", "start", "end", "name", "ID", "haplotype", "ship.name" )
starship_list = starships$ship.name

genome = "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/CEA10/A1163_edit6_liftoff.gff"
af_anno <- rtracklayer::import(con = genome,format = 'gff')
af_anno$type <- gsub("mRNA", "transcript", af_anno$type)

mcols(af_anno)$transcript_id = af_anno$locus_tag
mcols(af_anno)$gene_id = af_anno$locus_tag

path="C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/bigWig/"

atac=paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/FinalBigWig/", "151_ATAC_CEA10_WT_Hyphae37_marked.min30.max200.bin15.smooth45.bw")
  atac_bw <- load_BigWig(file_path=atac, file_name="ATACseq")
  
path2="C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/bigWig/"
k4=paste0(path2, "120-2_ChIP_WT_H3K4me2_Rep1.bin_25.smooth_75Bulk.bw")
  k4_bw <- load_BigWig(file_path=k4, file_name="H3K4me2")
k9=paste0(path2, "120-3_ChIP_WT_H3K9me3_Rep1.bin_25.smooth_75Bulk.bw")
  k9_bw <- load_BigWig(file_path=k9, file_name="H3K9me3")

  ##rerunning a select few
#  starship_list = c("navis04.var10", "navis07.var31", "navis10.Loki.var35")
#  ship="CEA10_s00019"
  
  starship_list
#  ship="Nebuchadnezzar.h1"
  
for (ship in starship_list) {
  print(ship)
  ship_coord = starships %>% filter(ship.name == ship)
  
  #set variables for start, end, and seqnames
  seq=ship_coord$seqnames
  start=ship_coord$start
  end=ship_coord$end
  
  print(c(ship, seq, start, end))

  ship_region = paste0(seq, ":", start, "-", end) 
  print(ship_region)
  ship_region=as(ship_region, "GRanges")
  
  ship_window = GRanges(seqnames = seq, ranges=IRanges(start = start, end = end))
  ship_overlap <- findOverlaps(af_anno, ship_window)
  ship_overlap_subset <- af_anno[queryHits(ship_overlap)]
  
  ext = 3000



#  atac_k4_bw <- c(atac, k4)
#  atac_k4_tracks <- load_BigWig(file_path=atac_k4_bw, file_name=c("ATAC-seq", "H3K4me2"))
  
  t_atac = coverage_vis_region(atac_bw, ship_region, up_extend = ext, down_extend = ext, y_lim = 30, style = "style_2", col_pal="darkorchid") + labs(title=ship) +
     theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),                                     
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(fill=NA, color= NA),
        panel.border=element_rect(fill=NA, color="black"),
        plot.margin = unit(c(0,0,0,0), "cm"),
        plot.title = element_text(hjust = 0.5))
        
t <- ggplot(t_atac)
  
    t_atac <- highlight_region(gg_object = t_atac, start = start, end = end)

  
  t_k4 = coverage_vis_region(k4_bw, ship_region, up_extend = ext, down_extend = ext, y_lim = 30, style = "style_2", col_pal="hotpink2") +
     theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),                                                                                                            
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(fill=NA, color= NA),
        panel.border=element_rect(fill=NA, color="black"),
        plot.margin = unit(c(0,0,0,0), "cm"))
  
    t_k4 <- highlight_region(gg_object = t_k4, start = start, end = end)

  t_k9 = coverage_vis_region(k9_bw, ship_region, up_extend = ext, down_extend = ext, y_lim = 60, style="style_2", col_pal="#3A5FCD") + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(fill=NA, color= NA),
        panel.border=element_rect(fill=NA, color="black"),
        plot.margin = unit(c(0,0,0,0), "cm"))

    t_k9 <- highlight_region(gg_object = t_k9, start = start, end = end)
  
    t_genes = transcript_vis_region(gene_anno = af_anno,region = ship_region, up_extend = ext, down_extend = ext, display_by = 'gene',show_name = NULL, display_mode = "squish") + 
      theme(axis.text.x= element_blank(), axis.ticks.x= element_blank(), axis.title.y = element_blank(), panel.border=element_blank(), panel.background = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"))

    t_k4_dim = get_dim(t_k4)
    set_dim(t_atac, t_k4_dim)
    set_dim(t_k9, t_k4_dim)
#  all <- t_atac_highlight + plot_spacer() + t_k4_highlight + plot_spacer() + t_k9_highlight + t_genes + plot_layout(guides = "collect", axes = "collect",  axis_titles = "collect", ncol = 1, heights = c(0.5, -.176, 0.5, -.2, 0.5, 0.2))
custom_spacer <- plot_spacer()+
                  theme(plot.margin = unit(c(0, 0, 0, 0),"lines"))

 all2 <- t_atac + custom_spacer + t_k4 + custom_spacer + t_k9 + t_genes + plot_layout(ncol=1, widths=c(1), heights=c(0.75, -0.047, 0.75, -0.047, .75, 0.08), guides="collect") 
 print(all2)
 assign(paste0("track_", ship), all2)

filename=paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/Plots/", "track_",ship,".png")
ggsave(filename=filename, plot=all2, device="png", width=5, height=2.5)


}
  custom_spacer <- plot_spacer()+
                  theme(plot.margin = unit(c(0.001, 0.001, 0.001, 0.001),"lines"))

  final_patch <- wrap_plots(track_Gnosis.h1, track_Janus.h2.2, track_Logos.h1, track_Nebuchadnezzar.h1, ncol=2, guides="collect")

  ggsave(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/Plots/final_patch.png", plot=final_patch, device="png", width = 9, height = 5)


\
  `      ``


```{r}
##Gnosis H1 has an outlier; plotting manually
  ship="Gnosis.h1"
  gnosis1_coord = starships %>% filter(ship.name == ship)
  
  #set variables for start, end, and seqnames
  seq=gnosis1_coord$seqnames
  start=gnosis1_coord$start
  end=gnosis1_coord$end
  

  gnosis1_region = paste0(seq, ":", start, "-", end) 
  print(gnosis1_region)
  gnosis1_region=as(gnosis1_region, "GRanges")
  
  gnosis1_window = GRanges(seqnames = seq, ranges=IRanges(start = start, end = end))
  gnosis1_overlap <- findOverlaps(af_anno, gnosis1_window)
  gnosis1_overlap_subset <- af_anno[queryHits(gnosis1_overlap)]
  
  ext = 3000
  
  atac_k4_bw <- c(atac, k4)
  atac_k4_tracks <- load_BigWig(file_path=atac_k4_bw, file_name=c("ATAC-seq", "H3K4me2"))
  
  t_atac_k4 = coverage_vis_region(atac_k4_tracks, gnosis1_region, up_extend = ext, down_extend = ext, y_lim=20, col_pal=c("darkorchid","hotpink2"))  +      
    theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                              axis.title.x = element_blank(),
                                       axis.text.y = element_text(size=10),
                                       axis.title.y = element_blank())
  
    t_atac_k4 <- highlight_region(gg_object = t_atac_k4, start = start, end = end)

#  t_atac = coverage_vis_region(atac_bw, gnosis1_region, up_extend = ext, down_extend = ext, col_pal="darkorchid")  + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y= element_text(6))
#    t_atac_highlight <- highlight_region(gg_object = t_atac, start = start, end = end)

#    t_k4 = coverage_vis_region(k4_bw, gnosis1_region, up_extend = ext, down_extend = ext, col_pal="hotpink2")  + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.title.x = element_blank(), axis.text.y= element_text(6))
#    t_k4_highlight <- highlight_region(gg_object = t_k4, start = start, end = end)

  t_k9 = coverage_vis_region(k9_bw, gnosis1_region, up_extend = ext, down_extend = ext, col_pal="#3A5FCD") +
    theme(axis.ticks.x = element_blank(),                                                                                                                                          axis.text.x = element_blank(),                                                                                                              axis.title.x = element_blank(),
                                       axis.text.y = element_text(size=10),
                                       axis.title.y = element_blank())
    t_k9 <- highlight_region(gg_object = t_k9, start = start, end = end)
    
  
    t_genes = transcript_vis_region(gene_anno = af_anno,region = gnosis1_region, up_extend = ext, down_extend = ext, display_by = 'gene',show_name = NULL, display_mode = "squish") + theme(axis.title.y= element_blank())

#  all <- t_atac_highlight + plot_spacer() + t_k4_highlight + plot_spacer() + t_k9_highlight + t_genes + plot_layout(guides = "collect", axes = "collect",  axis_titles = "collect", ncol = 1, heights = c(0.5, -.176, 0.5, -.2, 0.5, 0.2))

  

 all2 <- labs(title=ship) + t_atac_k4 + plot_spacer() + t_k9 + t_genes + plot_layout(ncol=1, heights=(c(1.5, -.25, .75, 0.2))) + theme(plot.title = element_text(hjust = 0.5))
 print(all2)
 assign(paste0("track_", ship), all2)

filename=paste0("C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/Plots/", "track_",ship,".png")
ggsave(filename=filename, plot=all2, device="png")
```


```{r}
## no k9 starships
no_k9 = c("Janus.h2.1", "Nebuchadnezzar.h1", "Osiris.h3", "Tardis.h1")
k9_ships = 
  
  
nok9_patch <- wrap_plots(track_Tardis.h1, track_Osiris.h3, track_Nebuchadnezzar.h1, track_Janus.h2.1, ncol=2, guides="collect")

k9_patch <- wrap_plots(track_navis04.var10, track_navis07.var31, track_Logos.h1, track_Gnosis.h2, track_navis10.Loki.var35, track_Logos.h2,   track_Gnosis.h1,track_Janus.h2.2, ncol=3, guides="collect")

ggsave(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/Plots/no_k9_patch.png", plot=nok9_patch, device="png", width = 9, height = 5)

ggsave(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/Plots/k9_patch.png", plot=k9_patch, device="png", width = 12, height = 5)


nok9_patch
##rename a few variable
  
  
assign("track_navis10.Loki.var35", track_navis10(Loki).var35)

k9_ships = gList(c(track_navis04.var10,track_navis07.var31, track_Logos.h1, track_Logos.h2))

patch1 <- wrap_plots(track_navis04.var10,track_navis07.var31, track_Logos.h1, track_Logos.h2, ncol=2, guides = "collect")
patch1
patch1 <- 
  wrap_plots(track_Tardis.h1, track_navis04.var10, track_navis07.var31, track_Osiris.h3, track_Nebuchadnezzar.h1, track_Logos.h1, ncol=1, guides = "collect", axes = "collect", axis_titles = "collect")
ggsave(filename="C:/Users/abbyd/Documents/Research/Bioinformatics/F32_PreliminaryData/Plots/patchtest.png", plot=patch1, device="png")


             patch1
             track_Gnosis.h2,track_Janus.h2.1, track_navis10.Loki.var35, track_Logos.h2, track_Gnosis.h1,track_Janus.h2.2, ncol=1, heights= c(10, 10, 10, 10, 1, 1))

  patch1
  Janus.h2.2_track / Osiris.h3_track
print(Janus.h2.2_track / Osiris.h3_track, plot_layout(ncol=1, heights=c(1,1)))
  
  
```


```{r}
echo


overlap <- findOverlaps(af_anno, window)
starship_subset <- af_anno[queryHits(overlap)]


  starship_subset <- subset(af_anno, seqnames(af_anno) == seqnames(window) &
                              start(af_anno) >= start(window) &
                              end(af_anno) <= end(window))
  
windows <- GRanges(seqnames = starships$V1, ranges=IRanges(start= starships$V2, end = starships$V3))

overlaps=findOverlaps(af_anno, windows)

# Extract the subset
subset_gr <- af_anno[queryHits(overlaps)]

bw = c("C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/BW_Downsampled/151_N17_ATAC_H1mnGreen_37conidia_Rep2_bin15_smooth45_min30.max200_Bulk.bw", "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/MappingOutput/BW_Downsampled/151-N1_ATAC_CEA10_WT__Rep1_marked.12M.bin15_smooth45_Bulk.bw")
genome = "C:/Users/abbyd/Documents/Research/Bioinformatics/AfATAC/Af_Genome/CEA10/A1163_edit6_liftoff.gff"

af_anno <- rtracklayer::import(con = genome,format = 'gff')
af_anno1 <- 
bigwig = load_BigWig(file_path = bw, file_name = c("Conidia", "Hyphae"))
t = coverage_vis_region(bigwig, starship, up_extend = 1000, down_extend = 1000)
p = t + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.title.x = element_blank())
t2 = transcript_vis_region(gene_anno = af_anno,region = starship, up_extend = 1000, down_extend = 1000, display_by = 'gene',show_name = 'gene_id', display_mode = "squish")
p2 = t2 + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.title.x = element_blank())
print(p + p2 + plot_layout(ncol = 1, heights = c(0.5,0.05)))

af_anno$type <- gsub("mRNA", "transcript", af_anno$type)

mcols(af_anno)$transcript_id = af_anno$locus_tag
mcols(af_anno)$gene_id = af_anno$locus_tag

setMethod("$", "GRanges", function(x, name) { # {{{
  elementMetadata(x)[, name]
}) # }}}

setMethod("$<-", "GRanges", function(x, name, value) { # {{{
  elementMetadata(x)[[ name ]] <- value
  return(x)
}) # }}}


```

